\begin{acknowledgements}

This thesis represents the accumulation of the work that I have done in the past four years. None of this would have been possible without the support, guidance and encouragement I have received from many people, before and during my PhD. To all of you, I am truly thankful.

Firstly, I would like to acknowledge the STFC for providing the funding to allow me to pursue my ambitions. This includes the two terrific years spent out at CERN. And to the Imperial HEP group for accepting me within their research group for, what I believe, is a successful PhD.

I would like to thank my supervisor, Sarah Malik, for her guidance throughout my PhD and, in particular, her patience and encouragement pushing me further. Although the outcome may differ from our original plan, I still believe it is an exceptional result that would not have been possible without you.

I would also like to thank the CMS collaborators and colleagues I have worked closely with throughout the past four years. Although this thesis does not include the work I performed on the hardware trigger or the $\alpha_{\mathrm{T}}$ analysis, I am still grateful to both the Calo layer 2 and $\alpha_{\mathrm{T}}$ groups. In particular, I would like to thank Alex Tapper for his support and guidance which resulted in my work being published early on during my PhD. I would also like to thank Rob Bainbridge for his continual support during the height of the $\alpha_{\mathrm{T}}$ analysis, and after. And to Nick Wardle, thank you for the countless discussions that has helped my understanding of statistics, as well as our more general discussions.

Thank you to all my friends, old and new, who I have relied heavily upon not only with my work, but in keeping me grounded and helping me through the stressful times. I could not wish for a better group of friends.

To my Mum, thank you for all the support and encouragement throughout my whole life. You have been inspirational and have always believed in me. And thank you to my whole family; my dad, step-dad and siblings, for your support and interest in my work.
    
\end{acknowledgements}