\chapter{Conclusions}
\label{chap:summary}

%\chapterquote{}{}

A measurement of the \PZ invisible width is performed with $\SI{35.9}{fb^{-1}}$ of data collected by the CMS experiment during the 2016 proton-proton collisions at $\sqrt{s}=\SI{13}{TeV}$. The invisible width is measured from the \metplusjets signature left in the detector as the \PZ boson decays into neutrinos which pass through undetected. These are compared to the muon and electron (collectively $\ell$) reference decays of the \PZ boson by collecting events with the \diellplusjets signature. The selection placed on these events is similar to the \metplusjets region to restrict to similar kinematics. In particular, the \recoil is used in place of the missing transverse energy as a proxy for the boson \pt in \IVj events. The modelling of this variable, as a proxy for the boson \pt, is validated in data by measuring its scale and resolution in events with a fully reconstructed dilepton for the boson \pt. This parameter is further used in the \ellplusjets control regions, where a data-driven estimate of the large \IWlvj background in the \metplusjets region is performed through the transfer factor approach.  Other backgrounds are significantly smaller and estimated from MC, apart from the QCD multijet process. A QCD multijet enriched sideband is collected by inverting the selection on \mindphi. Within this region, a correction factor for the QCD process is determined by extrapolation, with conservatively large uncertainties on the small background. In addition to these background estimations, the \IDYll process is divided into the \IZll and \Igstarll constituents to determine the contributions from each subprocess and their interference.

The invisible width is extracted with a simultaneous likelihood fit between the \metplusjets, \ellplusjets and \diellplusjets regions. The parameter of interest is extracted as a scaling parameter on the \IZvvj process, relative to \IZllj.  This parameter is translated into an invisible width measurement of
%
\begin{equation}
    \Gamma_{\mathrm{inv}} = 512 \pm 3\ (\mathrm{stat.})\ ^{+16}_{-15}\ (\mathrm{syst.})\ \mathrm{MeV}\ ,
\end{equation}
%
consistent with the SM and comparable to the LEP combined measurement. The measurement presented is systematically dominated, with the limiting sources from the jet energy scale, muon identification and trigger efficiency uncertainties.

The limiting systematic uncertainties imply that this measurements will not show significant gains from a larger dataset from the continual running of the LHC at increased luminosities. However, similar precision-based measurements are crucial tests of the SM, especially with the lack of BSM physics at the LHC.  These include furthering our understanding of the Higgs boson since its discovery at the LHC, as well as probing the electroweak sector with precision measurements.

The results shown are currently in the review process within the CMS collaboration. Most aspects of the analysis are in place, missing only the lepton momentum scale uncertainties. These uncertainties are expected to have a $1$--$\SI{2}{\%}$ affect on the final result, uncorrelated with all other uncertainties. In addition to these, a few techniques are under investigation to improve the limited MC sample size affecting the jet energy scale uncertainties with the hope of improving the $r_{\mathrm{inv}}$ measurement to about $\SI{2.7}{\%}$.